|

### 2021-05-27 article/article.Rmd
change intro text

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>. Corresponding author.

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>.

@@

Over the past decades, a large literature has examined the relationship between
legislature size and public expenditure. @weingast1981political provided the
general framework to analyse distributive politics. The authors argue that the
larger the number of legislative districts ($n$), the smaller the share of tax
burden each one will bear ($1/n$), thus legislators have an incentive to
overspend in their districts and transfer the costs to the entire polity. Early
studies that empirically tested the "law of $1/n$", as the theory is currently
known, indeed found a positive correlation between the number of legislature
seats and different measures of government spending, although their results
were mainly based on US state legislatures and the effect was often limited to
one house [e.g., @baqir2002districting; @gilligan1995deviations;
@gilligan2001fiscal].

@@

Since the wide range of empirical tests of the "law of $1/n$" has produced
conflicting results, scholars have expanded this research agenda and
investigated how institutional factors condition the original formulation of
the theory. For instance, authors such as @crowley2019law and
@pecorino2018supermajority accurately point out that collective action problems
have been overlooked in the literature, and recent findings indicate that
bicameralism [@maldonado2013legislatures], intergovernmental competition
[@crowley2015local], redistricting [@lee2018court], and party ideology
[@bjedov2014impact] strongly influence the relationship between seats and
spending. Moreover,

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to assess the impact of legislature size over
government spending across several dimensions. Out 
of the `r dim(dat)[1]` coefficients
included in our article sample, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Significant')])`\% of
them are positive and statistically significant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Insignificant')])`\%
are positive and statistically insignificant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Insignificant')])`\%
are negative and statistically insignificant, and `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Significant')])`\% are
negative and statistically significant. Although other studies aimed at solving this inconsistency, meta-analysis is the most appropriate tool, as it combines multiple outcomes into a single estimation and allows
us to draw robust conclusions from the aggregated data [@cooper2019handbook; @hedges1985statistical]. It can also identify potential
sources of study heterogeneity, enabling researchers to assess threats to
external validity and direct future efforts into more promising areas of
academic inquiry [@doucouliagos2008democracy].

We believe the scrutiny of the law expanded its initial scope in two directions. One group of scholars investigates the original institutional framework for the presence of coalitions, supermajorities, veto powers, and spending limits, which the theory disregards. They suggest these dynamics strongly influence the $1/n$ effect -- specifically in U.S. majoritarian local legislatures, for which the law was conceived. Another sector of the scholarship overrides this institutional context. They find that the different political arrangements in national congresses, State legislatures, regional European governments, etc. produce more nuanced results. For instance, intergovernmental competition, the size ratio between assemblies, and strong executives demonstrated the effect of legislative size on public spending could be non-linear, asymmetrical, and even negative. While these trends evolved, causal inference tools became more sophisticated, allowing this inquiry to grow also in roubstness. We address all of these developments in this study. We coded five different moderators that translate methodological advances in time, the similarity of the empirical cases to the original @weingast1981political framework, and key theoretical issues to identify sources of variation in the literature.

Aggregate results show that legislative size has no significant impact on
public spending. Our meta-analysis estimates are exactly half positive, and half negative. 
Meta-regressions suggest that our study sample is highly heterogeneous, and effects differ substantially according to study specifications. In the broadest aggregation models, unicameralism supports the positive effect predicted by the 
law of $1/n$. Since most unicameral cases in our sample refer to local governments (municipalities or districts), this result supports the theory strictly in its original institutional framework. It also indirectly endorses studies that find odd results in bicameralism. As follows, meta-regressions confirm previous findings that larger upper chambers spend more in terms of per capita expenditure than lower chambers [@gilligan1995deviations; @gilligan2001fiscal]. 
Moreover, non-majoritarian voting systems decrease government spending as a percentage of GDP, and they do not have a relevant impact on the other measures of public expenditure, following the idea that the $1/n$ effect grows weaker as the empirical cases distance from the original formulation of the law.

The meta-regressions also indicate that the $1/n$ effect found in earlier studies might be sensitive to more robust estimation methods. While papers that employ conventional linear regressions often find a positive relation between
legislature size and government budgets, recent studies using modern methods of
causal inference, specially those with regression discontinuity designs,
provide consistent evidence for a negative relation [@crowley2019law; @debenedetto2018effect; @hohmann2017effect; @lewis2019legislature]. This goes in contrast with
the original theory and supports the existence of a "reverse law of $1/n$". In summary, our results are in line 
with previous findings that lend partial support to the law of $1/n$ 
[@chen2007law; @crowley2015local; @primo2006stop]. An assessment of the effects of collective action costs in these analyses is crucial to determine the full mechanism operating behind these results. Overall, our analysis does not
provide robust support for the theory as originally conceived.

@@

Regarding the dependent variables included in the sample, `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='ExpPC')])` studies employ
public expenditure per capita, `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='logExpPC')])` papers use its natural logarithm, and `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='PCTGDP')])` of them
analyse the impact of legislature size on public expenditures as a percentage of
GDP. This indicates that the area has refined the original formulation of $1/n$
and tested the impact of larger legislatures on different measures of government
spending. Our independent variables are lower chamber size 
(`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='N')])`), the natural
logarithm of lower chamber size (`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='logN')])`), and upper chamber size (`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='K')])`). 
These six variables formed a $3 \times 3$ table, yet not all combinations were
available in the data. We found no studies that correlate public expenditure
per capita with either upper chamber size or the natural logarithm of lower
house size. Thus, our meta-analysis contains seven of the nine possible
variable combinations.

>>>>

[^danilo]: Senior Lecturer, School of Social and Political Sciences, University of Lincoln, <danilofreire@gmail.com>, <https://danilofreire.github.io>.

[^umberto]: Visiting Assistant Professor, Quantitative Theory and Methods Department, Emory University, <umberto.mignozzetti@emory.edu>, <http://umbertomig.com>. Corresponding author.

@@

Over the past decades, a large literature has examined the relationship between
legislature size and public expenditure. @weingast1981political provided the
general framework to analyse distributive politics. The authors argue that the
larger the number of legislative districts ($n$), the smaller the share of tax
burden each one will bear ($1/n$), thus legislators have an incentive to
overspend in their districts and transfer the costs to the entire polity. Early
studies that empirically tested the "law of $1/n$", as the theory is currently
known, indeed found a positive correlation between the number of legislature
seats and different measures of government spending, although these first
results were mainly based on US state legislatures and the effect was often
limited to one house [e.g., @baqir2002districting; @gilligan1995deviations;
@gilligan2001fiscal].

@@

Since many empirical tests of the "law of $1/n$" have produced conflicting
results, scholars have expanded this research agenda and investigated how
institutional factors condition the original formulation of the theory. For
instance, authors such as @crowley2019law and @pecorino2018supermajority
accurately point out that collective action problems have been overlooked in
the literature, and recent findings indicate that bicameralism
[@maldonado2013legislatures], intergovernmental competition
[@crowley2015local], redistricting [@lee2018court], and party ideology
[@bjedov2014impact] strongly influence the relationship between seats and
spending. Moreover, the literature has increasingly applied causal inference
designs to estimate the effect of the "law of $1/n$", and in contrast to
previous studies using panel data, regression discontinuity designs generally
indicate that more legislators decrease public expenditures
[@debenedetto2018effect; @hohmann2017effect; @lewis2019legislature]. In this
respect, scholars have long been aware of the theoretical and empirical
limitations of the "law of $1/n$", and the proliferation of new studies reflect
a conscious attempt to assess the robustness of the theory [@crowley2019law].

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to evaluate the impact of legislature size over
government spending across several dimensions. Our study sample mirrors the
diversity of the literature. Out of the `r dim(dat)[1]` coefficients
included in our main analysis, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Significant')])`\% of
them are positive and statistically significant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Insignificant')])`\%
are positive and statistically insignificant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Insignificant')])`\%
are negative and statistically insignificant, and `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Significant')])`\% are
negative and statistically significant. Meta-analysis provides a rigorous
approach to combine heterogeneous outcomes into a single estimation, and it
allows scholars to gain valuable insights from the aggregated data
[@cooper2019handbook; @hedges1985statistical]. Meta-analysis can also identify
potential sources of study variability, enabling researchers to assess threats
to external validity and direct future efforts into more promising areas of
academic inquiry [@doucouliagos2008democracy]. Research synthesis methods have
been successfully used in medicine and psychology since the 1970s
[@glass2015meta], and our work contributes to the burgeoning literature that
applies meta-analytic methods to understand challenging questions in political
science [@costa2017responsive; @doucouliagos2008democracy; @green2013field;
@lau2007effects; @schwarz2020supporting].

We run binomial tests and meta-regressions to further evaluate the relationship
between legislature size and government expenditures. We use the binomial tests
to asses whether the proportion of positive and negative coefficients in our
sample are statistically different from each other. In the meta-regressions, we
measure the effect of five moderators that capture different sources of
heterogeneity in the literature. As our sample includes both published and
unpublished articles, we also construct a series of funnel plots to test for
publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs (RDDs), we see that all
three papers included in our sample suggest that a higher number of legislators
leads to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature]. In this regard, it is possible that methodological
choices explain part of the disparity observed in the literature, as studies
which use causal inference methods consistently point to the same direction.
One limitation of this finding is that these studies only test the impact of
lower house size on the natural logarithm of expenditure per capita, and it
remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using our extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Since most unicameral
legislatures are located in local governments (municipalities or districts),
this result supports the theory strictly in its first formulation. Moreover,
our meta-regressions indicate that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests [@gilligan1995deviations; @gilligan2001fiscal]. Non-majoritarian
voting systems decrease government spending as a percentage of GDP, but they do
not have a relevant impact on the other measures of public expenditure,
following the idea that the $1/n$ effect grows weaker as the empirical cases
distance from the original definition of the law. Finally, meta-regression
results confirm that regression discontinuity models reduce public spending
estimates.

In summary, our results are in line with previous findings that lend partial
support to the law of $1/n$ [@chen2007law; @crowley2015local; @primo2006stop].
An assessment of the effects of collective action costs in these analyses is
crucial to determine the full mechanism operating behind these results.
Overall, our analysis does not provide robust support for the theory as
originally conceived.

@@

Regarding the dependent variables included in the sample, `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='ExpPC')])` studies employ
public expenditure per capita, `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='logExpPC')])` papers use its natural logarithm, and `r as.numeric(table(unique(dat[,c('id','depvar2')])$depvar2)[which(names(table(unique(dat[,c('id','depvar2')])$depvar2))=='PCTGDP')])` of them
analyse the impact of legislature size on public expenditures as a percentage of
GDP. This indicates that the area has refined the original definition of the "law of $1/n$"
and tested the impact of larger legislatures on different measures of government
spending. Our independent variables are lower chamber size 
(`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='N')])`), the natural
logarithm of lower chamber size (`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='logN')])`), and upper chamber size (`r as.numeric(table(dat$indepvar2)[which(names(table(dat$indepvar2))=='K')])`). 
These six variables formed a $3 \times 3$ table, yet not all combinations were
available in the data. We found no studies that correlate public expenditure
per capita with either upper chamber size or the natural logarithm of lower
house size. Thus, our meta-analysis contains seven of the nine possible
variable combinations.
